Using reference models in variable selection \vspace{.1in}

### 2022-02-21 reviews/rebuttal.tex

%\todo
%\begin{description}
%    \item [Mans] Nothing
%\end{description}

% Let's start point-by-point with Reviewer 1
\reviewersection
\begin{comment}
The paper is well written and organized an the topic analyzed is very relevant and practice. However, in my view, the goals of the paper, described in the introduction section, have already been reached in the excellent paper of Piironen et al. (2020) and, in this sense, the contributions of the current paper is very marginal. The framework of Piironen et al. (2020) is more general (they consider exponential family models and generalized linear models) and provide some theoretical results that, as they said “... help us to understand when the reference model could be helpful for parameter learning in linear submodels” (section 8 of Piironen et al., 2020). The only innovation that I can appreciate in the current paper is the proposal of an algorithm to solve a complete variable selection problem that, unfortunately, does not work very well as the authors explicitly recognize: “The great performance of projpred in minimal subset selection is not carried out over for the complete variable selection with iterative projpred ..., and the methods specifically designed for the complete variable selection perform better” (page 12, lines 24-27), even though they observe a better selection to iterative lasso. In my view, this is due to that the algorithm is too simple (as the authors explicitly recognize in the conclusions) and that the projective prediction approach was not designed for the complete variable selection. You should modify this algorithm in such a way that a more exhaustive exploration of the space of models was carried out, perhaps using Bayesian variable selection procedures based on the use of Occam’s window or something similar.
\end{comment}

\begin{reply}
	Thank you very much for your comments. We believe that in the first version of the paper, the purpose of our work was not clear and subject to possible misunderstandings. Our contribution takes clear inspiration from the work of \citet{paper:projpred}, but aims at a different goal: we studied the impact of using a reference model in general, disentangled from the specific variable selection algorithm.
	We now make our goals explicit already in Section 1, where we now write:

@@

	\vspace{0.5\baselineskip}

\end{reply}

\begin{point}
Please include the computational cost of compared approaches in Section 2.1. 
\end{point}

\begin{reply}
Thanks for the suggestion. We discuss the computational aspects in Section 2.2. In general, the reference model approach has the additional cost of fitting the reference model before performing selection. However, this reference models is usually available anyway when building up a model via robust workflows \citep{gelman2020bayesian}. Therefore, in these cases there is no additional cost. In any case we believe that the benefits of a more stable and efficient selection overcome the cost of fitting a reference model.
We added a new table (Table 3) reporting the computational time for the bodyf fat example, showing how the burden is divided between the fit of the reference model and the rest of the selection procedure.
\end{reply}

### 2022-02-21 paper/ref_approach.tex
final version for revision

\newcommand{\fede}[1]{\textcolor{red}{#1}}

@@

\fede{All these methods can be framed into the family of reference model approaches. 
The common denominator is the use of the reference model predictive information 
instead of simply observed data to guide the selection.}
Whatever the terminology or applied statistical framework, reference
models offer a powerful approach to improving variable selection, as we
will demonstrate in the present paper.

\fede{The goal of the present paper is to study the impact of reference model approaches by disentangling the benefit of using reference models per se from the benefit of specific variable selection algorithms. In particular, we:
\begin{itemize}
	\item propose a simple and intuitive approach to combine any variable selection method with the reference model approach. This allows us to investigate the benefit of using reference models independent of the specific variable selection method;
	\item perform extensive numerical experiments to compare variable selection methods with or without using a reference model, both for complete and minimal subset variable selection and assessing the quality of the selection;
	\item provide evidence supporting, in particular, the projection predictive approach as a principled way to use reference models in minimal subset variable selection.
\end{itemize}}

@@

The first two rows of Table \ref{tab:model_performances} show the
predictive performances, in terms of cross-validated root mean square
error (RMSE), of the full model and the selected models using projpred
or steplm.  There is no significant difference in predictive performance
of the selected models by different approaches, even if there is
clear difference in the number of selected variables. This can be explained by high
correlation between the variablesm and different combinations can
provide similar predictive accuracy.

@@

A properly designed reference model is able to filter parts of the
noise present in the data, and hence to provide an improved and more
stable selection process. This holds even if the reference model does
not perfectly resemble the true data generating process. 
\fede{Clearly, the reference model approach requires a sensible model. 
The construction of such a model should follow proper modelling workflow \citep[see, e.g.][]{gelman2020bayesian}. Better predictive models imply better selection results. 
The goodness of a reference model comes from its predictive ability which should be 
assessed via proper validation methods.} Our analyses indicate that the substantial 
reduction of variance attributable to noise is usually more important than small potential
bias due to model misspecification.
\fede{When the goal is ranking the variables for importance to
perform variable selection, it is known that large variance is
generally more harmful than small bias \citep{piironen2017comparison}.}

\fede{In general, there is no restriction on the type of model the reference models should be, and a sensible reference model does not even need to be Bayesian necessarily.
However, Bayesian methods can help in some of those situations
where MLE procedures struggle. 
It is known that as the number of parameters in the model increases, the MLE
estimator is dominated by shrinkage estimators 
\citep{stein1956inadmissibility, stein1961estimation, parmigiani2009decision, efron2011tweedie}.
The use of a prior in Bayesian inference automatically incorporates 
some kind of shrinkage in the Bayesian estimator under a given loss function \citep[see, e.g.][]{rockova2012hierarchical}. 
For the sake of our study, we rely on simple data structure which can be well described 
by linear regression models. However, more complex data can arise in practice, for example, including hierarchical structures. In these cases, MLE inference tends to become cumbersome, whereas the Bayesian framework provides a natural way to convey uncertainties and make inference through the joint posterior distribution of all parameters.}

@@

If one of the models to be compared is the full model, which can be
used as a reference model, there is no additional cost of using a
reference model as it was estimated as part of the analysis anyway.
Sometimes, including all the available variables in an
elaborate model can be computationally demanding.  In such a
case, even simpler screening or dimensionality reduction techniques,
as for example the supervised principal components
\citep{bair2006prediction,piironen2018}, can produce useful reference
models \citep{paper:projpred}. \fede{As an example, Table \ref{tab:computational_time} reports
the computational time for the body fat data example of Section \ref{bodyfat-1}. 
For reference model approaches, i.e. projpred and ref+step.lm, the main
time consuming operation is to fit the reference model. The burden of it depends clearly
on the specific application and type of model used.}

@@

In the body fat example above, our two simultaneous goals were to 
obtain good predictive performance and to select a smaller number of variables. 
When the goal is to select a minimal subset of variables,
which have similar predictive performance as the full model, we call it
\emph{minimal subset variable selection}. This minimal subset might
exclude variables which have some predictive information about the
target but, given the minimal subset, these variables are not able to
provide such additional information that would improve predictive
performance in a substantial manner.
The usual reason for this is that the relevant variables
which are not in the minimal subset are highly correlated with variables already in
the minimal subset. \fede{Such nature of the minimal subset makes the solution
not unique, except in the particular case of completely orthogonal predictors. 
The non-uniqueness of the solution is not a problem per se, as different samples of the data
are expected to give possible different minimal subsets but with same predictive power. 
However, it makes it difficult to define a proper
concept of stability of the selection.} 
We will return to a problem of finding all
the variables with some predictive power in Section~\ref{complete-selection}.
%
% The code to run all
% the experiments is available on GitHub\footnote{\url{https://github.com/fpavone/ref-approach-paper}}.

@@

In Figure \ref{fig:rmse_vs_fdr}, the predictive performance is shown
for different values of $n$ and $\rho$ in terms of RMSE and the false discovery rate (FDR, the ratio of the number of non-relevant selected
variables over the number of selected variables) of the selected submodel,
averaged after 100 data simulations. Use of a reference models greatly reduces the overfitting in the selection process, and thus produces much lower test set RMSE of the selected model for all the methods. projpred has the lowest FDR and RMSE. Using a reference model improves also FDR of steplm stepwise Bayesian liena rgegression significantly.

@@

\subsection{Reference model's effect and the projection predictive approach}
\fede{The previous experiments allowed us to study the impact of using reference models in variable selection and to disentangle their influence from that of the actual variable selection method, specifically in the minimal subset selection case.
The simulation study based on artificial data of Section 3.1 shows clear improvements in terms of predictive performance of the selected model and false discovery rate when the selection is based on a reference model, regardless of the actual variable selection method applied. These results are confirmed also in the real word data example with the body fat dataset. The stepwise selection achieves far better selection when coupled with the reference model approach (see Figure 5 and Table 4). However, the projection predictive approach remains the best method in any of the experiments we run and on all the performance indexes we measured. Although we designed a reference model approach for general selection method, the projection predictive approach is a principled and validated way to do the selection. Indeed, the purpose of the former is only for fair comparisons in our study rather than a ready-to-use selection method.}

@@

Projection predictive approach has been originally designed for minimal subset selection. For the comparison purposes, 
we modify the projection predictive approach for complete variable
selection by using it iteratively.  Applying the straightforward
implementation of projpred, we are able to select a minimal subset of
variables, which yield to a model with a predictive performance
comparable to the full model's predictive performance. 
 The iterative projection repeats
the projpred selection for different iterations, at each time
excluding the variables selected in the previous
iterations from the search.
 At each iteration, the selected submodel size corresponds to the
one having a predictive performance close enough to the baseline
model, which in this iterative version is the submodel with the
highest predictive score explored at the current iteration. This translates
in the following stopping rule at each iteration:
\begin{equation} 
\label{eq:rule_of_thumb}
\text{min} \{i\in \{0,\ldots,p\}: \text{P}(\text{elpd}_{i}-\text{elpd}_{\text{best}}>0)\geq \alpha \},
\end{equation}
where $i$ indexes the submodel size, ``best'' stands for the best predictive explored submodel
at the considered iteration, and the probability is computed from the normal approximation for the uncertainty in the cross-validation performance comparison \citep{Vehtari+etal:2017:practical,paper:projpred,Sivula+etal:2020:loo_uncertainty}. The algorithm terminates when the empty model (only intercept) satisfies
the stopping rule (see Algorithm \ref{alg:iterated_proj}).
The choice of the hyperparameter $\alpha$ is non-trivial, and we have 
observed sensitivity of the selection to such a choice, mainly when using cross-validation with a small
number of observations or not very predictive variables. We have used the stopping rule recommended by \citet{paper:projpred} for the minimal subset selection, but other stopping rules would be possible in the iterative case and worth a further research.
In our experiments, we chose the default value used in the \texttt{projpred} 
\texttt{R}-package, that is, $\alpha=0.16$. \fede{The choice of $\alpha$ to determine 
the submodel size is discussed in \cite{paper:projpred}. One possible way to proceed, 
but just as rule of thumb, is to choose $\alpha = 0.16$, which correspond to requiring 
that the submodel predictive score is at most at one standard deviation distance 
from the best submodel predictive score.}

>>>>

% \newcommand{\fede}[1]{\textcolor{red}{#1}}

@@

All these methods can be framed into the family of reference model approaches. 
The common denominator is the use of the reference model predictive information 
instead of simply observed data to guide the selection.
Whatever the terminology or applied statistical framework, reference
models offer a powerful approach to improving variable selection, as we
will demonstrate in the present paper.

The goal of the present paper is to study the impact of reference model approaches by disentangling the benefit of using reference models per se from the benefit of specific variable selection algorithms. In particular, we:
\begin{itemize}
	\item propose a simple and intuitive approach to combine any variable selection method with the reference model approach. This allows us to investigate the benefit of using reference models independent of the specific variable selection method;
	\item perform extensive numerical experiments to compare variable selection methods with or without using a reference model, both for complete and minimal subset variable selection and assessing the quality of the selection;
	\item provide evidence supporting, in particular, the projection predictive approach as a principled way to use reference models in minimal subset variable selection.
\end{itemize}

@@

The first two rows of Table \ref{tab:model_performances} show the
predictive performances, in terms of cross-validated root mean square
error (RMSE), of the full model and the selected models using projpred
or steplm.  There is no significant difference in predictive performance
of the selected models by different approaches, even if there is
clear difference in the number of selected variables. This can be explained by high
correlation between the variables and different combinations can
provide similar predictive accuracy.

@@

A properly designed reference model is able to filter parts of the
noise present in the data, and hence to provide an improved and more
stable selection process. This holds even if the reference model does
not perfectly resemble the true data generating process. 
Clearly, the reference model approach requires a sensible model. 
The construction of such a model should follow proper modelling workflow \citep[see, e.g.][]{gelman2020bayesian}. Better predictive models imply better selection results. 
The goodness of a reference model comes from its predictive ability which should be 
assessed via proper validation methods. Our analyses indicate that the substantial 
reduction of variance attributable to noise is usually more important than small potential
bias due to model misspecification.
When the goal is ranking the variables for importance to
perform variable selection, it is known that large variance is
generally more harmful than small bias \citep{piironen2017comparison}.

In general, there is no restriction on the type of model the reference models should be, and a sensible reference model does not even need to be Bayesian necessarily.
However, Bayesian methods can help in some of those situations
where MLE procedures struggle. 
It is known that as the number of parameters in the model increases, the MLE
estimator is dominated by shrinkage estimators 
\citep{stein1956inadmissibility, stein1961estimation, parmigiani2009decision, efron2011tweedie}.
The use of a prior in Bayesian inference automatically incorporates 
some kind of shrinkage in the Bayesian estimator under a given loss function \citep[see, e.g.][]{rockova2012hierarchical}. 
For the sake of our study, we rely on simple data structure which can be well described 
by linear regression models. However, more complex data can arise in practice, for example, including hierarchical structures. In these cases, MLE inference tends to become cumbersome, whereas the Bayesian framework provides a natural way to convey uncertainties and make inference through the joint posterior distribution of all parameters.

@@

If one of the models to be compared is the full model, which can be
used as a reference model, there is no additional cost of using a
reference model as it was estimated as part of the analysis anyway.
Sometimes, including all the available variables in an
elaborate model can be computationally demanding.  In such a
case, even simpler screening or dimensionality reduction techniques,
as for example the supervised principal components
\citep{bair2006prediction,piironen2018}, can produce useful reference
models \citep{paper:projpred}. As an example, Table \ref{tab:computational_time} reports
the computational time for the body fat data example of Section \ref{bodyfat-1}. 
For reference model approaches, i.e. projpred and ref+step.lm, the main
time consuming operation is to fit the reference model. The burden of it depends clearly
on the specific application and type of model used.

@@

In the body fat example above, our two simultaneous goals were to 
obtain good predictive performance and to select a smaller number of variables. 
When the goal is to select a minimal subset of variables,
which have similar predictive performance as the full model, we call it
\emph{minimal subset variable selection}. This minimal subset might
exclude variables which have some predictive information about the
target but, given the minimal subset, these variables are not able to
provide such additional information that would improve predictive
performance in a substantial manner.
The usual reason for this is that the relevant variables
which are not in the minimal subset are highly correlated with variables already in
the minimal subset. Such nature of the minimal subset makes the solution
not unique, except in the particular case of completely orthogonal predictors. 
The non-uniqueness of the solution is not a problem per se, as different samples of the data
are expected to give possible different minimal subsets but with same predictive power. 
However, it makes it difficult to define a proper
concept of stability of the selection.
We will return to a problem of finding all
the variables with some predictive power in Section~\ref{complete-selection}.
%
% The code to run all
% the experiments is available on GitHub\footnote{\url{https://github.com/fpavone/ref-approach-paper}}.

@@

In Figure \ref{fig:rmse_vs_fdr}, the predictive performance is shown
for different values of $n$ and $\rho$ in terms of RMSE and the false discovery rate (FDR, the ratio of the number of non-relevant selected
variables over the number of selected variables) of the selected submodel,
averaged after 100 data simulations. Use of a reference models greatly reduces the overfitting in the selection process, and thus produces much lower test set RMSE of the selected model for all the methods. projpred has the lowest FDR and RMSE. Using a reference model improves also FDR of steplm stepwise Bayesian linear regression significantly.

@@

\subsection{Reference model's effect and the projection predictive approach}
The previous experiments allowed us to study the impact of using reference models in variable selection and to disentangle their influence from that of the actual variable selection method, specifically in the minimal subset selection case.
The simulation study based on artificial data of Section 3.1 shows clear improvements in terms of predictive performance of the selected model and false discovery rate when the selection is based on a reference model, regardless of the actual variable selection method applied. These results are confirmed also in the real word data example with the body fat dataset. The stepwise selection achieves far better selection when coupled with the reference model approach (see Figure 5 and Table 4). However, the projection predictive approach remains the best method in any of the experiments we run and on all the performance indexes we measured. Although we designed a reference model approach for general selection method, the projection predictive approach is a principled and validated way to do the selection. Indeed, the purpose of the former is only for fair comparisons in our study rather than a ready-to-use selection method.

@@

Projection predictive approach has been originally designed for minimal subset selection. For the comparison purposes, 
we modify the projection predictive approach for complete variable
selection by using it iteratively.  Applying the straightforward
implementation of projpred, we are able to select a minimal subset of
variables, which yield to a model with a predictive performance
comparable to the full model's predictive performance. 
 The iterative projection repeats
the projpred selection for different iterations, at each time
excluding the variables selected in the previous
iterations from the search.
 At each iteration, the selected submodel size corresponds to the
one having a predictive performance close enough to the baseline
model, which in this iterative version is the submodel with the
highest predictive score explored at the current iteration. This translates
in the following stopping rule at each iteration:
\begin{equation} 
\label{eq:rule_of_thumb}
\text{min} \{i\in \{0,\ldots,p\}: \text{P}(\text{elpd}_{i}-\text{elpd}_{\text{best}}>0)\geq \alpha \},
\end{equation}
where $i$ indexes the submodel size, ``best'' stands for the best predictive explored submodel
at the considered iteration, and the probability is computed from a normal approximation of the uncertainty in the cross-validation performance comparison using the mean and standard error of the elpd difference between reference and submodel \citep{Vehtari+etal:2017:practical,paper:projpred,Sivula+etal:2020:loo_uncertainty}. The algorithm terminates when the empty model (only intercept) satisfies
the stopping rule (see Algorithm \ref{alg:iterated_proj}).
The choice of the hyperparameter $\alpha$ is non-trivial, and we have 
observed sensitivity of the selection to such a choice, mainly when using cross-validation with a small
number of observations or not very predictive variables. We have used the stopping rule recommended by \citet{paper:projpred} for the minimal subset selection, but other stopping rules would be possible in the iterative case and may be worth further research.
In our experiments, we chose the default value used in the \texttt{projpred} 
\texttt{R}-package, that is, $\alpha=0.16$. The choice of $\alpha$ to determine 
the submodel size is discussed in \cite{paper:projpred}. One possible way to proceed, 
but just as rule of thumb, is to choose $\alpha = 0.16$, which correspond to requiring 
that the submodel predictive score is at most at one standard deviation distance 
from the best submodel predictive score.
